---
title: "Introduction to `fertplanspatial` R package"
output: github_document
references:
- id: guidelines2020
  title: Parte agronomica, norme generali, Disciplinare di Produzione Integrata della Regione Lazio - SQNPI
  author:
  - family: Assessorato Agricoltura, Promozione della Filiera e della Cultura del Cibo, Ambiente e Risorse Naturali
  URL: 'http://www.regione.lazio.it/rl_agricoltura/?vw=documentazioneDettaglio&id=52065'
  publisher: Regione Lazio
  type: report
  issued:
    year: 2020
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(data.table)
library(fertplanspatial)
```


## Description

The goal of the package is to define fertilization plans for the fields of a farm and spatialize them. Fertilization plans in the Lazio region territory have to follow the agronomic guidelines with specific attention to [attachment no. 2](http://www.regione.lazio.it/binary/rl_main/tbl_documenti/AGC_DD_G01782_24_02_2020_Allegato1.pdf "PDF file of the Attachment 2 of the guidelines") [@guidelines2020] and rely on R package `fertplan` to perform the raw computations.


## Installation

`fertplanspatial` is currently in active development and not yet on CRAN, it may be installed from this GitHub repository though:

```{r, eval = FALSE}
# Install devtools package if not yet present in R library
# install.packages("devtools")

devtools::install_github("mbask/fertplanspatial")
```


```{r, results='asis', echo = FALSE}
res <- knitr::knit_child('vignettes/fertplanspatial_usage.Rmd', quiet = TRUE)
cat(res, sep = '\n')
```

## References